---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Loading libraries

```{r, message=F}
library(dplyr)
library(lattice)
```

## Loading and preprocessing the data

Dataset is loaded into object `data`.

```{r}
unzip('activity.zip')
data = read.csv('activity.csv', stringsAsFactors = T) %>% 
  mutate(clock = sprintf('%02d:%02d', floor(interval/100), interval %% 100),
         time = as.Date(sprintf("%s %s", date ,clock), format="%Y-%m-%d %H:%M"),
         date = as.Date(date)) %>% tbl_df
data
```

## What is mean total number of steps taken per day?

```{r}
steps.by.day = data %>% 
  group_by(date) %>%
  summarise(
    total = sum(steps, na.rm=T) )
steps.by.day
```

```{r, fig.width=6, fig.height=5}
hist(steps.by.day[['total']], main='Total number of steps taken each day', xlab='steps')
```

Calculated mean and median for number of steps taken each day.

```{r}
steps.by.day %>% 
  summarise(
    mean = mean(total),
    median = median(total) )
```

## What is the average daily activity pattern?

```{r}
steps.by.interval = data %>% 
  group_by(interval) %>%
  summarise(
    mean = mean(steps, na.rm=T),
    median = median(steps, na.rm=T))
with(steps.by.interval, plot(interval, mean, type='l', ylab='Average setps'))
```

The maximum number of steps is obtained at

```{r}
subset(steps.by.interval, mean == max(mean))
```

## Imputing missing values

The table has `r sum(!complete.cases(data))` rows with missing data `NA`.

```{r}
table(complete.cases(data))
```

To impute the `r sum(!complete.cases(data))` row I going to use the median calculated in the corresponding 5 minute period.

```{r}
data.imp = left_join(data, steps.by.interval, by='interval') %>%
  mutate(
    steps = ifelse(is.na(steps), median , steps) ) %>%
  select_(.dots = names(data))
data.imp
```

The histogram

```{r}
steps.by.day.imp = data.imp %>% 
  group_by(date) %>%
  summarise(
    total = sum(steps, na.rm=T) )
```

```{r, fig.width=6, fig.height=5}
hist(steps.by.day.imp[['total']], main='Total number of steps taken each day', xlab='steps')
```

  * Before imputation
  
```{r}
steps.by.day %>% 
  summarise(
    mean = mean(total),
    median = median(total) )
```

  * After imputation
  
```{r}
steps.by.day.imp %>% 
  summarise(
    mean = mean(total),
    median = median(total) )
```

> After imputation we have seen that the estimate of the mean differs.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
data.imp = data.imp %>% 
  mutate(
    period = ifelse(weekdays(date) %in% c('Saturday', 'Sunday'), 'weekend', 'weekday') )
steps.by.interval_period.imp = data.imp %>% 
  group_by(interval, period) %>%
  summarise(
    mean = mean(steps, na.rm=T),
    median = median(steps, na.rm=T))
```

```{r, fig.width=6, fig.height=5}
xyplot(mean~interval|period, data=steps.by.interval_period.imp, type='l', 
       layout=c(1,2), ylab='Number of steps', xlab='Interval')
```